rbcL community comparison

### 2019-11-02 code/rbcl_community_compare.Rmd
combined UniFrac and rel abund into multi panel

```{r libs, include=F}
library(tidyverse)
library(here)
library(GUniFrac)
library(vegan)
library(ggrepel)
# library(vegalite)
# library(fossil)
library(picante)

@@

```{r input, include=F}
rbcl_rel_abund <- read_csv(here("data/clean/rbcl_rel_abund_clean.csv"))
idtax <- read_csv(here("data/clean/rbcl_taxonomy_clean.csv"))
field <- read_csv(here("data/clean/field_clean.csv"))

rbcl_asv_id_key <- read_csv(here("intermediate_output/rbcl_asv_key.csv"))
rbcl_asv_otu_key <- read_csv(here("intermediate_output/asv_otu_key.csv"))
alias <- read_csv(here("data/sample_aliases.csv"))
rbcl.tree <- read.tree(here("intermediate_output/tree/rbcl/asv613.fasta.treefile"))
```

```{r unifrac, include=F}
ra_wide <- rbcl_rel_abund %>% 
  select(asv_id, sample_id, rel_abund) %>% 
  pivot_wider(names_from = asv_id, values_from = rel_abund, values_fill = list(rel_abund = 0)) %>% 
  column_to_rownames("sample_id")

# root tree with Trebouxiophyceae
root <- idtax %>% 
  filter(class=="Trebouxiophyceae") %>% 
  head(1) %>% 
  pull(asv_id)

@@

@@

```{r unifrac_plot, echo=F}
# plot
unifrac_plot_data <- nmds %>% 
  scores() %>% 
  as_tibble(rownames = "sample_id") %>% 
  left_join(field, by="sample_id") %>% 
  left_join(alias, by="sample_id")

@@

```{r setup_rel_abund_barplot, fig.width=10, fig.height=7, include=F}
plot_data <- rbcl_rel_abund %>% 
  left_join(rbcl_asv_otu_key, by="asv_id") %>% 
  group_by(sample_id, otu_name) %>% 
  summarise(rel_abund = sum(rel_abund)) %>%   
  left_join(field, by="sample_id") %>%
  left_join(alias, by="sample_id") %>% 
  ungroup() %>% 
  mutate(rel_abund = if_else(rel_abund <0.01, 0.01, rel_abund)) # artificially boost rel abund <1% to 1% so its visible on plot

@@

```{r plot_barplot, fig.width=10, fig.height=7, echo=F}
cust_theme <- theme(legend.position = "none",
                    panel.grid.major.x = element_blank(), # remove vertical lines in background
                    panel.grid.minor.x = element_blank(),
                    axis.line=element_blank(), # remove x axis
                    axis.title.x=element_blank(), 
                    axis.text.x=element_blank(),
                    axis.ticks.x=element_blank(),
                    panel.spacing = unit(-0.01, "lines"), # remove space between facet panels
                    strip.text.x = element_text(angle = 90, hjust=1.0, vjust = 0.8), # rotate facet labels 90
                    strip.background = element_blank()) # remove grey boxes around facet labs

# order OTUs
otu_ord <- c("Sanguina A","Chloromonas I","Chloromonas G","Chloromonas muramotoi E","Chloromonas krienitzii D","Chloromonas spp. C","Chlainomonas B","Chlainomonas F","Algae H", "Trebouxiophyceae J")

plot <- ggplot(plot_data, aes(x=alias %>% fct_reorder(elev_m), y=rel_abund)) +
  geom_bar(aes(fill=otu_name), stat="identity") + # classic stacked relative abundance plot
  coord_flip() +
  facet_grid(~otu_name %>% fct_relevel(otu_ord), switch = "x") +
  labs( x="Sample ID (from lowest to highest elevation)") +
  cust_theme
plot
# add sparkline on side for elevation
```

@@

faiths <- re_wide_mat %>% 
  picante::pd(rbcl.tree) %>% 
  # attach elevation data
  rownames_to_column("sample_id") %>% 
  left_join(field, by="sample_id") %>% 
  left_join(alias, by="sample_id")

@@

```{r output, include=F}
# ggsave
```

>>>>

```{r, echo=FALSE}
knitr::opts_chunk$set(fig.path='figs/', echo=FALSE, warning=FALSE, message=FALSE)
```

```{r libs, include=F}
library(tidyverse)
library(here)
library(GUniFrac)
library(vegan)
library(ggrepel)
# library(vegalite)
# library(fossil)
library(picante)
library(gridExtra)

@@

```{r input, include=F}
rel_abund <- read_csv(here("data/03_filtered/rbcl/filtered_cleaned_rbcl_rel_abund.csv")) # not looking at the depth samples, controls, etc
tax <- read_csv(here("data/02_tidied/tidied.rbcl_taxonomy.csv"))
field <- read_csv(here("data/02_tidied/tidied_field.csv"))

asv_seqs <- read_csv(here("data/asv_seq_data/rbcl_asv_sequence_key.csv"))
otus <- read_csv(here("output/asv_otu_key.csv"))

rbcl.tree <- read.tree(here("output/tree/rbcl/all_asvs/rbcl_aligns_all.fasta.treefile")) ### this tree should only contain the asvs we found in the "main" sites, not looking at controls, etc
```

```{r unifrac, include=F}

ra_wide <- rel_abund %>% 
  select(sample_id, asv_id, rel_abund) %>% 
  arrange(asv_id) %>% 
  pivot_wider(names_from = asv_id, values_from = rel_abund, values_fill = list(rel_abund = 0)) %>% 
  column_to_rownames("sample_id")

# root tree with Trebouxiophyceae
root <- tax %>% 
  filter(class=="Trebouxiophyceae") %>% 
  head(1) %>% 
  pull(asv_id)

@@

# check for differences between asvs in tree and rel abund table (all names must appear in tree)
asvs_rel_abund <- colnames(ra_wide)
asvs_tree <- rbcl.tree$tip.label %>% sort()

length(asvs_rel_abund)
length(asvs_tree)

setdiff(asvs_rel_abund, asvs_tree) # output elts unique to asvs_rel_abund

@@

```{r unifrac_plot, echo=F}
# plot
unifrac_plot_data <- nmds %>% 
  scores() %>% 
  as_tibble(rownames = "sample_id") %>% 
  left_join(field, by="sample_id")

@@

```{r setup_rel_abund_barplot, include=F}
plot_data <- rel_abund %>% 
  left_join(otus, by="asv_id") %>% 
  group_by(sample_id, otu_name) %>% 
  summarise(rel_abund = sum(rel_abund)) %>%   
  left_join(field, by="sample_id") %>%
  ungroup() %>% 
  mutate(rel_abund = if_else(rel_abund <0.01, 0.01, rel_abund)) # artificially boost rel abund <1% to 1% so its visible on plot

@@

```{r plot_barplot, echo=F}
cust_theme <- theme(legend.position = "none",
                    panel.grid.major.x = element_blank(), # remove vertical lines in background
                    panel.grid.minor.x = element_blank(),
                    axis.line=element_blank(), # remove x axis
                    axis.title.x=element_blank(), 
                    axis.text.x=element_blank(),
                    axis.ticks.x=element_blank(),
                    panel.spacing = unit(-0.01, "lines"), # remove space between facet panels
                    strip.text.x = element_text(angle = 90, hjust=1.0, vjust = 0.8), # rotate facet labels 90
                    strip.background = element_blank()) # remove grey boxes around facet labs

otu_names %>% distinct(otu_name)

# order OTUs
otu_ord <- c("Chlainomonas A","Sanguina B", "Chloromonas spp. C","Chloromonas krienitzii D","Chloromonas muramotoi E","Chloromonas F", "Algae G","Chloromonas H", "Trebouxiophyceae I")

barplot <- ggplot(plot_data, aes(x=alias %>% fct_reorder(elev_m), y=rel_abund)) +
  geom_bar(aes(fill=otu_name), stat="identity") + # classic stacked relative abundance plot
  coord_flip() +
  facet_grid(~otu_name %>% fct_relevel(otu_ord), switch = "x") +
  labs( x="Sample ID") +
  cust_theme
barplot

```

```{r combine, fig.width=7, fig.height=10}
combined_fig_3 <- grid.arrange(uf_plot, barplot, nrow=2)
```

@@

faiths <- re_wide_mat %>% 
  picante::pd(rbcl.tree) %>% 
  # attach elevation data
  rownames_to_column("sample_id") %>% 
  left_join(field, by="sample_id")

@@

```{r output, include=F}
ggsave(filename = here("manuscript/figs/3_community/combined_fig_3.pdf"), plot=combined_fig_3, height=10, width=7)
ggsave(filename = here("manuscript/figs/supplementary/diversity.png"), plot = faith_date_plot, height = 5, width = 7)
```
